---
title: "R Notebook Neo4j Movie Graph Example for the neo4r package"
knit: (function(input_file, encoding) {
 out_dir <- 'docs';
 rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))
 } )
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
  html_notebook:
    code_folding: hide
    df_print: paged
    fig_height: 8
    fig_width: 11
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r results='hide', warning=FALSE, message=FALSE}

library(knitr)
library(neo4r)
library(tibble)
library(dplyr)
library(tidyverse)
library(visNetwork)
library(RColorBrewer)
library(kableExtra)
```
# Setup Neo4j Environment {.tabset .tabset-fade}
## Connect to Neo4j / Verify schema
```{r neo4j-connection}
neo4jServerURL <- "http://localhost:7474"
neo4jUser <- "neo4j"
neo4jPassword <- "admin"
con <- neo4j_api$new(url = neo4jServerURL, user = neo4jUser, password = neo4jPassword)
if (!is.null(con)) {
  if (file_test('-f', '/.dockerenv')) {  # CK THIS - verify if you're in a docker container
    print("Appear to be running in a docker container")
    Sys.sleep(3)
  }
  if (con$ping() == 200) {
    print(paste("Successfully connected to Neo4j version", con$get_version()))
    if (!is.na(Sys.getenv("RSTUDIO", unset = NA)) && interactive() ) {
      print("You are running in R Studio and can use Connections pane to browse schema.")
      launch_con_pane(con)
    }
  } else {
    cat(paste("Neo4j Server = ", neo4jServerURL, ", Neo4j User = ", neo4jUser, "\n"))
  }
}
```
## Validate movied DB
```{r neo4j-Movie-Validation }
requiredLabels <- tibble(labels = c("Person", "Movie"))
requiredRelationships <-
  tibble(labels = c("ACTED_IN", "DIRECTED", "PRODUCED", "WROTE","FOLLOWS","REVIEWED"))

label_matches <-
  nrow(inner_join(requiredLabels, con$get_labels(), by = 'labels')) - nrow(requiredLabels)
relationship_matches <-
  nrow(inner_join(requiredRelationships, con$get_relationships(), by = 'labels')) - nrow(requiredRelationships)

if (label_matches == 0 && relationship_matches == 0) {
  cat("Using database with data from the Neo4j movies graphgist\n")
} else {
  cat("Not using a database with the Movie Graph data!\nThe database is missing:\n")
  if (label_matches != 0) {
    cat("Node(s):\n")
    print.data.frame(anti_join(requiredLabels, con$get_labels(), by = 'labels'))
  }
  if (relationship_matches != 0) {
    cat("RELATIONSHIP(S):\n")
    print.data.frame(anti_join(requiredRelationships, con$get_relationships(), by = 'labels'))
  }
  # send_cypher('https://raw.githubusercontent.com/dfgitn4j/Rmovies/master/create_movies.cypher', con = con)
  # create movie graph database using the :play movies query
  # Change this to deep linking into desktop? 
  #   See https://medium.com/neo4j/neo4j-desktop-release-command-bar-security-deep-linking-be899d72b947 for how to use
  cat("... creating the Movie Graph example data (you will need the apoc libraries installed) ...")
  "CALL apoc.cypher.runFile('https://raw.githubusercontent.com/dfgitn4j/Rmovies/master/create_movies.cypher')" %>%
    call_neo4j(con)
}
```
## Utility Functions
Utility functions

*neo4r helper functions*

- flatten property structures
- unnest nodes & relationships

*visNetwork*

Setup data structures

- Assign consisent colors for each node
- Assign what to use for a label

*Initialize a visNetwork display variable*

- General graph display function for nodes and edges

### Neo4r Helper Functions
```{r visNetwork-neo4r-helper-functions}

# Node and relationship structure returned by Neo4j is constant.
# properties can be lists of lists
flattenProperties <- function(props) {
  prop_lists <- lapply(props, unlist, recursive=FALSE)    # get 1st level
  prop_names <- unique(unlist(lapply(prop_lists, names)))
  prop_lists <- lapply(prop_lists, function(x) {
    y <- as.list(x)[prop_names]  # get values
    names(y) <- prop_names       # add names
    lapply(y, function(z) if(is.null(z)) NA else z)})
  flatten_data <- lapply(prop_names, function(x) sapply(prop_lists, "[[", x))
  names(flatten_data) <- as.list(prop_names)
  data.frame(flatten_data, stringsAsFactors=FALSE)
}
  
unnest_nodes2 <- function(nodes, what = c("all", "label", "properties")) {
  what <- match.arg(what)
  id_df <- tibble(id = nodes$id) # node id's
  if (what %in% c('label', 'all') ) {
    labels <- cbind(nodes[,c('id','label')])
    if (what == 'label') { return(labels)}
  } 
  if (what %in% c('properties', 'all') ) {
    properties<-cbind(id_df, flattenProperties(nodes$properties))
    if (what == 'properties') { return(properties)}
  } 
  # what == 'all'
  return(inner_join(labels, properties, by = 'id') )
}

unnest_relationships2 <- function(relationships, what = c("relationships", "properties", "all")) { 
  what <- match.arg(what)
  if (is.null(relationships)) {
    edge_data <- NULL
  } else {
    id_df <- tibble(id = relationships$id)
    if (what %in% c('relationships', 'all') ) {
      rels<-select(relationships, id, startNode, endNode, type)
      if (what == 'relationships') { return(rels) }
    } 
    if (what %in% c('properties', 'all') ){ 
      props<-cbind(tibble(id = relationships$id), flattenProperties(relationships$properties))
      if (what == 'properties') { return(props) }
    }
    return(inner_join(rels, props, by = "id"))
  }
}

```
### Visualization properties
```{r graph-visualize-properties}
# set constant colors for :Labels by visNetwork label value
# put node count in actual_value
# add text field to tell what property field to use for node display
#  - is group (label) property name df
q<-"CALL db.labels() yield label 
      CALL apoc.cypher.run('match (:'+label+') return count(*) as count', null) YIELD value
      RETURN label, value.count as count"  %>% call_neo4j(con, type="row")

nodeVisuals<<-cbind.data.frame( q$label,
                                color=colorRampPalette(brewer.pal(11,"Set3"))(nrow(q$label)), 
                                stringsAsFactors = FALSE )
nodeVisuals <- rename(nodeVisuals, label = value)  # change name from query name value to labels
nodeVisuals$group<-nodeVisuals$label
nodeVisuals$actual_value<-q$count$value
nodeVisuals<-select(nodeVisuals, everything()) %>% 
  left_join(tibble('group' = c('Movie', 'Person', 'Test'), 'text' = c('title',  'name', 'group')), by = 'group' )
```
###Table output format functions
```{r tableOutput-utilities}
# use kable for nice formatting of table output
kableTable <- function(qry_res) {
   # test to see if list of mutiple values
  if ( reduce( (lapply(qry_res, function(x) { nrow(x)})), sum ) > 1 ) 
    qry_res<-(map(qry_res, "value")) # return columns labeld value
  qry_res %>%  # get rid of boolean 
  as.data.frame() %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")
}

```
### visNetwork utiity functions
```{r visNetwork-utilities}
# set groups for a visNetwork object based on labels
setGroups <- function(g_disp, n_visuals, shape="circle") {
  for (gr in n_visuals$label) {
    g_disp<-visGroups(g_disp, groupname = gr, 
                      color = n_visuals[n_visuals$label == gr,]$color,
                      shape=shape,
                      borderwidth=2)
  }
  return(g_disp)
}

# rename startNode, endNode, and type for visNetwork
visNetEdgeDisplayProps <- function (edge_df) {

  edge_df <- rename(edge_df, from = startNode, to = endNode, label = type)
  edge_df$shadow = TRUE
  edge_df$labelHighlightBold = TRUE
  edge_df$arrows.to.scaleFactor = 2
  edge_df$arrows = "to"
  
  return(edge_df)
}

# adding color consistency and joining nodeVisuals for layout
visNetworkNodeDisplayProps <- function(node_df, use_props) {
  node_df<-select(node_df, one_of(c('id', 'label'))) %>% 
    unnest() %>% 
    left_join(nodeVisuals, by = c('label' = 'group')) %>% 
    rename( group = label) %>%
    inner_join(node_df, by = 'id') %>% 
    group_by(id) %>% mutate(gid = row_number()) %>% # add row nbr by grouped id's to take first label if there's more than one
    filter(gid == 1) %>%
    ungroup()  # grouping will add id column for select queries if left intact

    #  using a property for display text as joined from nodeVisuals above, or group (Label) if there is not one defined
  if (use_props) { 
    # label column is display text for nodes PUT IN is.na() test to return a default
    node_df$text<-replace_na(node_df$text, 'group') # use group (neo4j :Label) if no text is defined to display in node
    node_df$label<-apply(select_at(node_df, node_df$text), 1, function(x) na.omit(x)[1])  
  } else { # use :Label for node text, e.g. Movie
    node_df$label<-node_df$group
  }
  node_df$shadow<-TRUE
  node_df$shape<-'circle'
  
  return(node_df)
}
# initialize a visNetwork structure
setInitialGraph <- function(qry_graph_output, title = 'Graph', use_properties_for_label=FALSE) {
  
  node_data<<-unnest_nodes2(qry_graph_output$nodes)
  edge_data<<-unnest_relationships2(qry_graph_output$relationships)
  # node_data<<-unnest_nodes(qry_graph_output$nodes)
  # edge_data<<-unnest_relationships(qry_graph_output$relationships)

  node_data<-visNetworkNodeDisplayProps(node_data, use_properties_for_label)
  edge_data<-visNetEdgeDisplayProps(edge_data)
  
  return (graph_disp<-visNetwork(node_data, edge_data, width = "100%", main = title) %>%
            visLayout(randomSeed = 100))
}

```
# Graph metadata {.tabset .tabset-fade}

## Show colors assigned to node labels

Display sizes nodes based on number of nodes in the graph
```{r}
# simple graph to display the consistent colors used for Neo4j node :Label(s) 
# w

disp_nodes <- nodeVisuals

disp_nodes$shape<-'dot'  # circle would put label in node, but ignore value

disp_nodes$size<-ifelse(disp_nodes$actual_value < 5, 5, disp_nodes$actual_value) # hard to hover w/ node size < 5
disp_nodes$title<- apply(disp_nodes, 1, function(x) { paste0("<p>", x[['group']], " (", toString(x[['actual_value']]), " Nodes )</p>") })
graph_disp<-visNetwork(disp_nodes, NULL, width = "100%", main = "Colors For Graph :Labels") 
graph_disp<-setGroups(graph_disp, nodeVisuals, shape = 'square')
graph_disp%>% visHierarchicalLayout(direction = "UD") %>% 
  visInteraction(dragNodes = FALSE, dragView = FALSE, zoomView = FALSE) %>%
  visLegend(zoom = FALSE) 
```

## Show movie graph structure 

```{r}
q<-"call db.schema()" %>% call_neo4j(con, type = 'graph') 
disp_graph<-setInitialGraph(q, title = "Graph Metadata", use_properties_for_label = FALSE) 
disp_graph %>% visPhysics(solver = 'barnesHut', 
              barnesHut =  list(gravitationalConstant = -60000, centralGravity = 1, springLength = 150, springConstant = 0.05, avoidOverlap = 1))

```

# Movie Graph 

Pop-cultural connections between actors and movies

The Movie Graph is a mini graph application containing actors and directors that are related through the movies they've collaborated on.

This guide will show you how to:

1. Find: retrieve individual movies and actors
2. Query: discover related actors and directors
3. Solve: the Bacon Path

## Find {.tabset .tabset-fade}
### Find the actor named "Tom Hanks"...

`MATCH (tom {name: 'Tom Hanks'}) RETURN tom`
```{r find-tom-hanks}
res<-"MATCH (tom {name: 'Tom Hanks'}) RETURN tom" %>% call_neo4j(con, type = 'row') 
kableTable(res)
#! NOT SURE why test column
```
### Find the movie with title "Cloud Atlas"...

`MATCH (cloudAtlas {title: 'Cloud Atlas'}) RETURN cloudAtlas`
```{r find-cloud-atlas}
res<-"MATCH (cloudAtlas {title: 'Cloud Atlas'}) RETURN cloudAtlas" %>% call_neo4j(con, type = 'row') 
kableTable(res)
```
### Find 10 people...

`MATCH (people:Person) RETURN people.name,  people.born LIMIT 10`
```{r find-10-people}
res<-"MATCH (people:Person) RETURN people.name,  people.born LIMIT 10" %>% call_neo4j(con, type = 'row') 
kableTable(res)
```
### Find movies released in the 1990s...

`MATCH (nineties:Movie) WHERE nineties.released >= 1990 AND nineties.released < 2000 RETURN nineties.title`
```{r 1990s-movies}
res<-"MATCH (nineties:Movie) WHERE nineties.released >= 1990 AND nineties.released < 2000 RETURN nineties.title" %>% call_neo4j(con, type = 'row') 
kableTable(res)
```
## Query {.tabset .tabset-fade}
Finding patterns within the graph.

1. Actors are people who acted in movies
2. Directors are people who directed a movie
3. What other relationships exist?

### List all Tom Hanks movies...

`MATCH p=(tom:Person {name: 'Tom Hanks'})-[:ACTED_IN]->(tomHanksMovies) RETURN p`
```{r list-tom-hanks-movies}
# original query does not return all the info needed for a visualization (automatically done by desktop)
res<-"MATCH p=(tom:Person {name: 'Tom Hanks'})-[:ACTED_IN]->(tomHanksMovies) RETURN p" %>% call_neo4j(con, type = 'graph') 
disp_graph<-setInitialGraph(res, title="Tom Hanks Movies", use_properties_for_label = TRUE)
disp_graph %>% visPhysics(enabled = FALSE, solver = 'forceAtlas2Based', 
           forceAtlas2Based = list(avoidOverlap = 0.75, gravitationalConstant=-100))
```
### Who directed "Cloud Atlas"?

`MATCH (cloudAtlas {title: 'Cloud Atlas'})<-[:DIRECTED]-(directors) RETURN directors.name`
```{r directed-cloud-atlas}
res<-"MATCH (cloudAtlas {title: 'Cloud Atlas'})<-[:DIRECTED]-(directors) RETURN directors.name" %>% call_neo4j(con, type = 'row') 
kableTable(res)
```
### Tom Hanks' co-actors...

`MATCH (tom:Person {name: 'Tom Hanks'})-[:ACTED_IN]->(m)<-[:ACTED_IN]-(coActors) RETURN coActors.name`
```{r tom-hanks-co-actors}
res<-"MATCH (tom:Person {name: 'Tom Hanks'})-[:ACTED_IN]->(m)<-[:ACTED_IN]-(coActors) RETURN coActors.name" %>% 
  call_neo4j(con, type = 'row') 
kableTable(res)
```
### How people are related to "Cloud Atlas"...

`MATCH (people:Person)-[relatedTo]-(:Movie {title: 'Cloud Atlas'}) RETURN people.name, Type(relatedTo), relatedTo`

```{r related-to-cloud-atlas}
#! DOES NOT WORK 
# call_neo4j(con, type = 'row') below gives "Error in cbind_all(x) : Argument 3 must be length 1, not 4""
res<-"MATCH (people:Person)-[relatedTo]-(:Movie {title: 'Cloud Atlas'}) RETURN people.name, Type(relatedTo), relatedTo" %>% 
  call_neo4j(con, type = 'graph') 
disp_graph<-setInitialGraph(res, title="People related to Cloud Atlas", use_properties_for_label = TRUE)
disp_graph %>% visPhysics(enabled = FALSE, solver = 'forceAtlas2Based', 
           forceAtlas2Based = list(avoidOverlap = 0.75, gravitationalConstant=-10000))
```

## Solve {.tabset .tabset-fade}
You've heard of the classic "Six Degrees of Kevin Bacon"? That is simply a shortest path query called the "Bacon Path".

1. Variable length patterns
2. Built-in shortestPath() algorithm

### Movies and actors up to 4 "hops" away from Kevin Bacon

`MATCH p=(bacon:Person {name:'Kevin Bacon'})-[*1..4]-(hollywood) RETURN DISTINCT p`
```{r 6-degrees-kevinB}
res<-"MATCH p=(bacon:Person {name:'Kevin Bacon'})-[*1..4]-(hollywood) RETURN DISTINCT p"  %>% call_neo4j(con, type = 'graph')
disp_graph<-setInitialGraph(res, title="Four 'Hops' From Kevin Bacon", use_properties_for_label = TRUE)
where_is_kevinB<-match('Kevin Bacon', disp_graph$x$nodes$label)
disp_graph$x$nodes$shape[where_is_kevinB]<-'box'
disp_graph$x$nodes$color[where_is_kevinB]<-'#FF4141'
disp_graph %>% visPhysics(enabled = FALSE, solver = 'forceAtlas2Based', 
           forceAtlas2Based = list(avoidOverlap = 0.75, gravitationalConstant=-10000)) %>%
  visOptions(highlightNearest = TRUE, selectedBy = "name") %>%
  visInteraction(navigationButtons = TRUE)
```

### Bacon path, the shortest path of any relationships to Meg Ryan

`MATCH p=shortestPath(
  (bacon:Person {name:'Kevin Bacon'})-[*]-(meg:Person {name:'Meg Ryan'}))RETURN p`
```{r shortest-path-kevin-meg}
# original query does not return all the info needed for a visualization (automatically done by desktop)
res<-"MATCH p=shortestPath(
  (bacon:Person {name:'Kevin Bacon'})-[*]-(meg:Person {name:'Meg Ryan'}))RETURN p" %>% call_neo4j(con, type = 'graph') 
disp_graph<-setInitialGraph(res, title="Shortest Path Between Kevin and Meg", use_properties_for_label = TRUE)
where_is_kevinB<-match('Kevin Bacon', disp_graph$x$nodes$label)
disp_graph$x$nodes$shape[where_is_kevinB]<-'box'
disp_graph$x$nodes$color[where_is_kevinB]<-'#FF4141'
disp_graph %>% visPhysics(enabled = FALSE, solver = 'forceAtlas2Based', 
           forceAtlas2Based = list(avoidOverlap = 0.75, gravitationalConstant=-10000))
```
